Progress Report

### 2023-09-22 Docs/ProgressReport.Rmd
add GLMM adaptive model summary
re-run data application part and record time

- 80% (7017) subjects for model fitting, 20% (1753) subjects for out-of-sample prediction
- 3.3 minutes spent on model fitting, 20 minutes spent on prediction (about 0.01 per subject)
- Three subjects had failed approximation, all happened with the shortest observed track
- We can either assume that the latent function value is constant within the same bin, or use linear interpolation to fill in the gaps between bin midpoints. Here I wanna compare if this choice would make any different. Also, there are a log of other methods to fill in the gaps, such as projection. 
- For linear interpolation, I set the edges to be the closest data extreme.

@@

# interpolation for each subject
for(i in unique(df_test_full$id)){
  # interpolation t540
  interp_i_t540 <- approx(x = df_test[df_test$id==i, "bin"], 
                     y = df_test[df_test$id==i, "pred_t540"],
                     xout = 541:J, rule = 2)
  df_test_full$pred_t540_interp[df_test_full$id==i & df_test_full$sind>540] <- interp_i_t540$y

@@

  # interpolation t1020
  interp_i_t1020 <- approx(x = df_test[df_test$id==i, "bin"], 
                     y = df_test[df_test$id==i, "pred_t1020"],
                     xout = 1021:J, rule = 2)
  df_test_full$pred_t1020_interp[df_test_full$id==i & df_test_full$sind>1020] <- interp_i_t1020$y
}

@@

Let's look at four subjects as an example:

@@

@@

```
- Subjects with numeric issue

```{r, fig.height=6, fig.width=9}
df %>% filter(id %in% skip_id) %>% 
  ggplot()+
  geom_point(aes(x=sind, y=Y), size = 0.2)+
  facet_wrap(~id)+
  geom_vline(xintercept = c(540))
```

- Calculated AUC

@@

- when fitting model with **random intercept + slope** on the same full training set (even after scaling covariates): Error: vector memory exhausted (limit reached?). 
- when fitting model with **only random intercept** on the same full training set: Error regarding a large coefficient value. One fix is to re-scale covariates. So I re-scaled minute index by dividing them by J, so the range is within (0, 1], and tried again. This fitting procedure took 20.72 minutes to finish
- Out-of-sample prediction took 12.18 minutes. About 0.007 minute per subject.
- I'd also exclude the subjects with failed Laplace approximation here.

@@

@@

@@

@@

@@

colnames(tb_auc) <- rep(c("9am", "1pm", "5pm"), 3)

tb_auc %>%
  kable(digit = 4) %>%
  kable_styling(full_width = F) %>% 
  add_header_above(c("Time spent on model fitting" = 1, "3.3" = 6, "20.72" = 3), bold = F) %>%
  add_header_above(c("Time spent on prediction" = 1, "20.0" = 6, "12.2" = 3), bold = F) %>% 
  add_header_above(c(" "=1, "fGFPCA"=3, "fGFPCA+interpolation" = 3, "GLMMadaptive"=3))
```

- fGFPCA almost always have higher AUC. It at least is as good as GLMMadaptive.
- Computationally, fGFPCA is about 10 minute faster (33%).

>>>>

- 80% (7017) subjects for model fitting, 20% (1753) subjects for out-of-sample prediction
- Used PIRLS for local GLMMs step

@@

# interpolation for each subject
t1 <- Sys.time()
for(i in unique(df_test_full$id)){
  # interpolation t540
  interp_i_t540 <- approx(x = df_test[df_test$id==i, "bin"], 
                     y = df_test[df_test$id==i, "pred_t540"],
                     xout = 541:J, rule = 2)
  df_test_full$pred_t540_interp[df_test_full$id==i & df_test_full$sind>540] <- interp_i_t540$y

@@

  # interpolation t1020
  interp_i_t1020 <- approx(x = df_test[df_test$id==i, "bin"], 
                     y = df_test[df_test$id==i, "pred_t1020"],
                     xout = 1021:J, rule = 2)
  df_test_full$pred_t1020_interp[df_test_full$id==i & df_test_full$sind>1020] <- interp_i_t1020$y
}
t2 <- Sys.time()
t_interp <- t2-t1

@@

- `r round(t_local_glmm, 2)` minutes spent on fitting local GLMMs, `r round(t_fpca, 2)` seconds on FPCA model, `r round(t_pred, 2)` minutes on out-of-sample prediction, `r round(t_interp, 2)` minutes on interpolation to the original grid

- Let's look at four subjects as an example:

@@

### A closer look at interpolation

- We can either assume that the latent function value is constant within the same bin, or use linear interpolation to fill in the gaps between bin midpoints. Here I wanna compare if this choice would make any different. Also, there are a lot of other methods to fill in the gaps, such as projection. 
- For linear interpolation, I set the edges to be the closest data extreme.
- The interpolation itself spent `r round(t_interp, 2)` minutes.

@@

### Numeric problems

- Three subjects had failed approximation, all happened with the shortest observed track
- Does it have to do with still the length of observed track? Or the arbitrary choices in the binning procedure (bin width).

```{r, fig.height=3, fig.width=9}
df %>% filter(id %in% skip_id) %>% 
  ggplot()+
  geom_point(aes(x=sind, y=Y), size = 0.2)+
  facet_wrap(~id)+
  geom_vline(xintercept = c(540))
```

<!-- Calculate AUC -->

@@

- when fitting model with **random intercept + slope** on the same full training set (even after scaling covariates): Error: vector memory exhausted (limit reached?). 
- when fitting model with **only random intercept** on the same full training set: Error regarding a large coefficient value. One fix is to re-scale covariates. So I re-scaled minute index by dividing them by J, so the range is within (0, 1], and tried again. This fitting procedure took 20.72 minutes to finish

@@

- `r round(t_est_adglmm, 2)` minutes on model fitting and `r round(t_pred_adglmm, 2)` minutes on out-of-sample prediction.

- Below is the model fit on the trainign set

```{r}
summary(adglmm_mod)
```

@@

- Below is an example of predicted probablity on four random subjects

@@

<!-- Calculate AUC  -->

@@

- To use the same sample for comparison, I'd exclude the subjects with failed Laplace approximation from the test set for both methods

@@

colnames(tb_auc) <- rep(c("9am", "1pm", "5pm"), 3)
```

```{r}
# time
t_fit <- c(t_local_glmm+t_fpca, t_local_glmm+t_fpca, t_est_adglmm)
units(t_fit) <- "mins"

t_test <- c(t_pred, t_pred+t_interp, t_pred_adglmm)
# units(t_pred)

t_total <- t_fit+t_test

idx1 <- idx2 <- idx3 <- c(1, 3, 3, 3)
names(idx1) <- c("Total time", as.character(round(t_total, 2)))
names(idx2) <- c("Time on prediction", as.character(round(t_test, 2)))
names(idx3) <- c("Time on model fitting", as.character(round(t_fit, 2)))

t_fit <- as.character(round(t_fit, 2))
t_test <- as.character(round(t_test, 2))
t_total <- as.character(round(t_total, 2))

```{r}
tb_auc %>%
  kable(digit = 4, table.attr = "style = \"color: black;\"") %>%
  kable_styling(full_width = F) %>% 
  add_header_above(header=idx1) %>%
  add_header_above(header=idx2) %>%
  add_header_above(header=idx3) %>%
  add_header_above(c(" "=1, "fGFPCA"=3, "fGFPCA+interpolation" = 3, "GLMMadaptive"=3))
```

- fGFPCA does not always have higher AUC, but definitely most of the time. 
- In fact there are two scenarios where GLMMadaptive overperforms fGFPCA slightly: 1) given 9am to predict 1-5pm; and 2) given 1pm and predict 5-9pm. Both of them have a 4-hour interval apart from the maximum observation time and and prediction window. Could it have anything to do the cyclic trend in the data? Is 4h the length of a cycle? I tried to look at it but really couldn't see anything...there are too many points lumped together!

```{r}
df %>% filter(id %in% rand_id2) %>%
  ggplot()+
  geom_point(aes(x=sind, y=Y), size = 0.2)+
  geom_vline(xintercept = c(540, 780, 1020, 1260))+
  scale_x_continuous(breaks = c(540, 780, 1020, 1260),
                     labels = c("9am", "1pm", "5pm", "9pm"))+
  facet_wrap(~id)
```
